---
title: "Plot"
author: "Evangeline Reynolds"
date: "2/13/2019"
output: html_document
---

# Now for some plotting.

We may use some of the wrangline steps, and feed the wrangled data into a ggplot() with a pipe:  %>%.





---
# Country grouped trend plot faceted by continent




---

```{r complete_01, eval=F, echo=F, comment = " "}
gapminder %>%
  mutate(gdp_billions = 
           gdpPercap * 
           pop/1000000000) %>% 
  ggplot() +
  aes(x = year) + 
  aes(y = gdp_billions) +
  geom_line() +
  aes(group = country) +
  scale_y_log10() +
  aes(col = continent) +
  facet_wrap( ~ continent) +
  scale_color_discrete(guide = F) +
  theme_minimal()
```

`r apply_reveal("complete_01")`



---

# Scatterplot with smoothing

---

```{r complete_02, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 1997) %>%
  ggplot() +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  scale_x_log10() +
  geom_smooth()
```

`r apply_reveal("complete_02")`


---

# Faceted scatterplot

---

```{r complete_02_1, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 1997) %>%
  ggplot() +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  scale_x_log10() +
  geom_point() +
  facet_wrap(~ continent) +
  aes(size = pop) +
  labs(size = "population")
```

`r apply_reveal("complete_02_1")`





---

# bar plot (count for each category)

---

```{r complete_04, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 2007) %>%
  ggplot() +
  aes(x = continent) +
  geom_bar()
```

`r apply_reveal("complete_04")`



---

# column plot (specifying height of bars)

---

```{r complete_05, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  tally() %>%
  ggplot() +
  aes(x = continent, y = n) +
  geom_col()
```

`r apply_reveal("complete_05")`



---

# box plot

---

```{r complete_06, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 2007) %>%
  ggplot() +
  aes(x = continent, y = lifeExp) +
  geom_boxplot() +
  geom_jitter(height = 0, width = .2) +
  stat_summary(fun.y = mean, 
               geom = "point", 
               col = "goldenrod3", 
               size = 5)
```

`r apply_reveal("complete_06")`



---

# violin + jitter plot

---

```{r complete_06_1, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 2007) %>%
  ggplot() +
  aes(x = continent, y = lifeExp) +
  geom_violin() +
  geom_jitter(height = 0, width = .2)
```

`r apply_reveal("complete_06_1")`


---

# Wrangle plot

---

```{r complete_07, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 2007) %>%
  ggplot() +
  aes(x = lifeExp) +
  geom_histogram() +
  aes(fill = continent) +
  scale_fill_viridis_d() +
  facet_wrap(~continent) +
  geom_rug()

```

`r apply_reveal("complete_07")`



---

# Overlapping histograms

---

```{r complete_08, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 2007, 
         continent == "Africa" | 
           continent == "Americas") %>%
  ggplot() +
  aes(x = lifeExp) +
  aes(fill = continent) +
  geom_histogram(alpha = .2, 
                 position = "identity") +
  theme_light() +
  geom_rug() +
  aes(col = continent)

```

`r apply_reveal("complete_08")`



---

# Line graph

---

```{r complete_08_1, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(country == "Argentina") %>%
  ggplot() +
  aes(x = year, y = lifeExp) +
  geom_point() +
  geom_line() +
  aes(alpha = year) +
  aes(col = year) +
  scale_color_viridis_c() +
  theme_classic()

```

`r apply_reveal("complete_08_1")`


---

# Line graph variant

---

```{r complete_09, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(country == "Argentina" | 
           country == "Chile") %>%
  ggplot() +
  aes(x = year, y = lifeExp) +
  geom_point() +
  geom_line() +
  aes(group = country) +
  aes(col = country) +
  aes(size = pop)

```

`r apply_reveal("complete_09")`




---

# Jitter plot, two-way categorical data

---

```{r complete_10, eval=F, echo=F, comment = " "}
gapminder %>%
  mutate(life_exp_70_plus = 
           lifeExp > 70) %>%
  filter(year == 1997) %>%
  ggplot() +
  aes(x = continent, 
      y = life_exp_70_plus) +
  geom_jitter(width = .25, 
              height = .25) +
  aes(col = continent) +
  scale_color_discrete(guide = F) +
  theme_dark()
```

`r apply_reveal("complete_10")`







---

# Some ggplot food for thought with questions.

---

# Plot 1

---

```{r plot1, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 2007) %>%
  ggplot() +
  aes(x = continent) +
  geom_bar() +
  aes(col = continent) +
  aes(fill = continent) +
  scale_fill_viridis_d() +
  coord_flip() +
  geom_bar(width = .5, fill = "magenta") +
  aes(fill = NULL) +
  labs(x = "") +
  labs(title = "My plot title!")
```


`r apply_reveal("plot1")`

---

# Questions, Plot 1

How many geometric layers ("geoms") are in the final plot?

To which geometric layer do the aesthetics mapping "fill" apply, ie represents a variable in the data?

What does coord_flip() do?

How is NULL used?

When adding a geom layer, where does it appear, behind or in front of other layers?



---
# Plot 2

---

```{r plot2, eval=F, echo=F, comment = " "}
gapminder %>%
  filter(year == 2007) %>%
  ggplot() +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point(alpha = .5) +
  geom_rug(size = 1) +
  aes(col = continent) +
  aes(col = lifeExp) +
  scale_x_log10() +
  aes(size = gdpPercap) +
  aes(size = pop) +
  geom_point(col = "darkgreen", size = 1) +
  facet_wrap(~ continent)
```


`r apply_reveal("plot2")`


---
# Questions Plot 2

How many geometric layers are in the final plot?

What does geom_rug do?

There are two times when the aesthetic mapping col is set and two times when the aesthetic mapping size is set. Which is the final setting?



---
# Plot 3

---

```{r plot3, eval=F, echo=F, comment = " ", message = FALSE, warning=F}
gapminder %>%
  filter(continent == "Europe") %>%
  ggplot() +
  aes(x = year) +
  aes(y = lifeExp) +
  aes(alpha = year) +
  geom_point() +
  geom_line() +
  aes(col = country) +
  aes(col = NULL) +
  aes(group = country)
```


`r apply_reveal("plot3")`


---
# Questions Plot 3

How many geometric layers are in the final plot?

What does the alpha setting do?




